Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities

### 2020-09-26 writing/transect_paper.Rmd
use ggplot instead of tikz to draw amplicon map, and put together subfigures using ggarrange instead of latex

---
title: "Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities"
author: 
  - Brendan Furneaux^1,\*^,
  - Mohammad Bahram^2,3^,
  - Anna Rosling^4^,
  - Nourou S. Yorou^5^,
  - Martin Ryberg^1^
date: |
      | ^1^Program in Systematic Biology,
        Department of Organismal Biology,
        Uppsala University,
        Uppsala, Sweden
      | ^2^Department of Ecology,
        Swedish University of Agricultural Sciences,
        Uppsala, Sweden
      | ^3^Institute of Ecology and Earth Sciences,
        University of Tartu,
        Tartu, Estonia
      | ^4^Program in Evolutionary Biology,
        Department of Ecology and Genetics,
        Uppsala University,
        Uppsala, Sweden
      | ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
        LEB,
        University of Parakou,
        Parakou, Benin
      | ^\*^Corresponding author, `brendan.furneaux@ebc.uu.se`
bibliography:
  - "all.bib"
  - "R.bib"
biblio-style: apa
always_allow_html: true
output:
  bookdown::pdf_document2:
    template: bare_template.tex
    fig_crop: true
    keep_tex: true
    toc: false
    latex_engine: lualatex
    includes:
      in_header: preamble.tex
    citation_package: biblatex
    suppress-bibliography: true
    extra_dependencies: "subfig"
  bookdown::word_document2: 
    keep_md: yes
fontsize: 12pt
geometry: ["left=2.5cm", "right=2.5cm", "top=3cm", "bottom=3cm"]
graphics: yes
---

@@

(ref:region-asv) **(a)** Partial map of rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA (thick lines), and internally transcribed spacer (ITS) regions (thin black lines).  D1-3 (grey) represent the first three variable regions in LSU, while LSU1-4 (black) represent the conserved regions. Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces. **(b)** Total number of DADA2 ASVs vs. fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from a set of long PacBio amplicon sequences using `LSUx`. Shorter and more conserved regions yield a greater fraction of successfully mapped reads<!-- This sentence is a result shown in the figure (although it is also theoretically expected)-->. At a given fraction of mapped reads, more variable regions yield a greater number of unique ASVs<!-- This sentence is how to interpret the figure -->.

```{r regions, fig.cap="(ref:region-asv-s). (ref:region-asv)", fig.scap="(ref:region-asv-s)", fig.subcap=rep("", 2), fig.ncol = 1, fig.align = "center"}
knitr::include_graphics("rDNA_amplicons", auto_pdf = TRUE)
readd(region_table) %>%
  dplyr::filter(seq_run == "pb_500") %>%
  mutate(reads = reads / sum(filter(readd(readcounts), seq_run == "pb_500", step == "demux")$nreads)) %>%
  dplyr::mutate_at("region", fct_recode, "5.8S" = "5_8S") %>%
  ggplot(aes(reads, ASVs, color = length_med, label = region)) +
  geom_point() +
  ggrepel::geom_text_repel() +
  scale_color_gradientn(colors = c("blue", "red", "green1"), limits = c(30, 2000), breaks = c(50, 150, 400, 1000), trans = "log10", name = "median length") +
  scale_x_continuous(name = "Fraction of reads mapped",limits = c(0, 1)) +
  ylab("Unique ASVs")
```

>>>>

---
title: "Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities"
author: 
  - Brendan Furneaux^1,\*^,
  - Mohammad Bahram^2,3^,
  - Anna Rosling^4^,
  - Nourou S. Yorou^5^,
  - Martin Ryberg^1^
date: |
      | ^1^Program in Systematic Biology,
        Department of Organismal Biology,
        Uppsala University,
        Uppsala, Sweden
      | ^2^Department of Ecology,
        Swedish University of Agricultural Sciences,
        Uppsala, Sweden
      | ^3^Institute of Ecology and Earth Sciences,
        University of Tartu,
        Tartu, Estonia
      | ^4^Program in Evolutionary Biology,
        Department of Ecology and Genetics,
        Uppsala University,
        Uppsala, Sweden
      | ^5^Research Unit in Tropical Mycology and Plant-Fungi Interactions,
        LEB,
        University of Parakou,
        Parakou, Benin
      | ^\*^Corresponding author, `brendan.furneaux@ebc.uu.se`
bibliography:
  - "all.bib"
  - "R.bib"
biblio-style: apa
always_allow_html: true
output:
  bookdown::pdf_document2:
    template: bare_template.tex
    fig_crop: true
    keep_tex: true
    toc: false
    latex_engine: lualatex
    includes:
      in_header: preamble.tex
    citation_package: biblatex
    suppress-bibliography: true
  bookdown::word_document2: 
    keep_md: yes
fontsize: 12pt
geometry: ["left=2.5cm", "right=2.5cm", "top=3cm", "bottom=3cm"]
graphics: yes
---

@@

(ref:region-asv) **(a)** Partial map of rDNA showing the 5.8S rDNA, partial SSU and LSU rDNA, and internally transcribed spacer (ITS) regions.  D1-3 represent the first three variable regions in LSU, while LSU1-4 represent the conserved regions. Primer sites used in this study are indicated in red (forward primers) and blue (reverse primers), and the resulting amplicons are shown with green braces. **(b)** Total number of DADA2 ASVs vs. fraction of demultiplexed reads successfully mapped to ASVs for different rDNA regions extracted from a set of long PacBio amplicon sequences using `LSUx`. Shorter and more conserved regions yield a greater fraction of successfully mapped reads<!-- This sentence is a result shown in the figure (although it is also theoretically expected)-->. At a given fraction of mapped reads, more variable regions yield a greater number of unique ASVs<!-- This sentence is how to interpret the figure -->.

```{r regions, fig.cap="(ref:region-asv-s). (ref:region-asv)", fig.scap="(ref:region-asv-s)", fig.align = "center"}
ggpubr::ggarrange(
  readd(amplicon_plot),
  readd(length_diversity_plot),
  ncol = 1,
  labels = "auto",
  heights = c(0.6, 1)
)
```
